### 2025-02-08 ESC195 Lecture Notes/main.tex
feb 8

\subsection{Comparison Tests}

@@

>>>>

\subsection{Other Comparison Tests}

@@

\begin{theorem}[Limit Comparison Test]
    Let $\{a_n\}$ and $\{b_n\}$ be two sequences such that $a_n, b_n > 0$ for all $n$. Then:
        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = c > 0$, then either both series $\sum a_n$ and $\sum b_n$ converge or both diverge.
        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = 0$ and $\sum b_n$ converges, then $\sum a_n$ converges.
        \item If $\lim_{n \to \infty} \frac{a_n}{b_n} = \infty$ and $\sum b_n$ diverges, then $\sum a_n$ diverges.
    We can prove the first part of the theorem by epsilon-delta proof. 
    \\
    Proof of part 1: Let $\lim_{n \to \infty} \frac{a_n}{b_n} = c > 0$. Then, for all $\epsilon > 0$, for all $n > N$, we have:
        |\frac{a_n}{b_n} - c| &< \epsilon \\
        \intertext{Choose $\epsilon = \frac{c}{2}$, then we have:}
        |\frac{a_n}{b_n} - c| &< \frac{c}{2} \\
        \frac{c}{2} < \frac{a_n}{b_n} &< \frac{3c}{2} \\
    We can then use the basic comparison test to show that the series converge or diverge.
    Let $a_n = \frac{1}{n^3 - n}$ and choose $b_n = \frac{1}{n^3}$. We can use the limit comparison test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \frac{n^3}{n^3 - n} \\
    Since $\sum b_n$ converges, then $\sum a_n$ converges.
    Let $a_n = \frac{\ln(n/1000)}{n}$ and chooise $b_n = \frac{1}{n}$. We can use the limit comparison test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \frac{a_n}{b_n} &= \lim_{n \to \infty} \frac{\ln(n/1000)}{n} \\
        &\to \infty
    Since $\sum b_n$ diverges, then $\sum a_n$ diverges.
\subsection{Alternating Series and Absolute Convergence}
\begin{definition}[Alternating Series]
    An alternating series is a series of the form:
        \sum_{n=1}^{\infty} (-1)^{n-1} a_n = a_1 - a_2 + a_3 - a_4 + \ldots
    where $a_n > 0$ for all $n$.
    The proof for this is straight forward. We can show that the series converges by showing that the sequence $\{S_n\}$ is bounded above and below. Left as an exercise to the reader (Hint: Consider both the odd and even partial sums).
    Let $a_n = \frac{1}{n}$, then the series $\sum (-1)^{n-1} \frac{1}{n}$ is an alternating series.
\paragraph{Note} $\sum \cos n$ is not an alternating series, though it alternates between positive and negative values, the negative and postive values are not necessarily next to each other.
\begin{theorem}[Alternating Series Test]
    Let $\{a_n\}$ be a sequence such that $a_n > 0$ for all $n$ and $\lim_{n \to \infty} a_n = 0$. Then the alternating series $\sum (-1)^{n-1} a_n$ converges if:
        \item $a_{n+1} \le a_n$ for all $n$.
        \item $\lim_{n \to \infty} a_n = 0$.
\begin{theorem}[Value of Alternating Series]
    The value of an alternating series is given by:
        S = a_1 - a_2 + a_3 - a_4 + \ldots = \lim_{n \to \infty} S_n
    where $S_n$ is the $n$th partial sum of the series. Additionally, the limit would always be between two consecutive partial sums.
    Let $S = e^{-1} = \sum (-1)^{n-1} \frac{1}{n!}$, we can use the above to estimate the value of $e^{-1}$. For $n=5$ we have:
        e^{-1} &= 0.366 \pm 0.0015
\begin{definition}[Absolute Convergence]
    A series $\sum a_n$ is said to be absolutely convergent if the series $\sum |a_n|$ converges. Conversely, a series is said to be conditionally convergent if the series converges but the absolute value of the series diverges.
    Let $a_n = \frac{(-1)^{n-1}}{n}$, then the series $\sum \frac{(-1)^{n-1}}{n}$ is conditionally convergent (Hint: 1/2 $L$).
\begin{theorem}[Absolute Convergence Test]
    If the series $\sum |a_n|$ converges, then the series $\sum a_n$ converges.
    We can prove this by using the comparison test. We can show that $-|a_n| \le a_n \le |a_n|$ and $0 \le a_n + |a_n| \le 2|a_n|$. Thus, we can use the comparison test to show that the series converges. Since $\sum a_n + |a_n|$ converges, then $\sum a_n$ converges.
\subsection{Ratio and Root Tests}
\begin{theorem}[Root Test]
    Let $\{a_n\}$ be a sequence, and let $L = \lim_{n \to \infty} \sqrt[n]{|a_n|}$. Then:
        \item If $L < 1$, then the series $\sum a_n$ converges.
        \item If $L > 1$, then the series $\sum a_n$ diverges.
        \item If $L = 1$, then the test is inconclusive.
    Proof of part 1: Let $L < 1$. Then, choose $\mu$ such that $L < \mu < 1$. Then for some $N$, $\sqrt[n]{|a_n|} < \mu$ for all $n > N$, then $a_n < \mu^n$ for all $n > N$. We can then use the comparison test to show that the series converges since $\sum \mu^n$ converges for $|\mu| < 1$.
    Let $a_n = \left ( \frac{n^2 + 1}{2n^2 + 1} \right )^n$. We can use the root test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \sqrt[n]{\left ( \frac{n^2 + 1}{2n^2 + 1} \right )^n} &= \lim_{n \to \infty} \frac{n^2 + 1}{2n^2 + 1} \\
        &= \frac{1}{2}
    Let $a_n = \frac{2^n}{n}$. So $\sqrt[n]{\frac{2^n}{n}} = \frac{2}{n^{1/n}}$. So the limit $2$ (denom. converges to 2), is greater than 1, thus the series diverges.
\begin{theorem}[Ratio Test]
    Let $\{a_n\}$ be a sequence, and let $L = \lim_{n \to \infty} \left | \frac{a_{n+1}}{a_n} \right |$. Then:
        \item If $L < 1$, then the series $\sum a_n$ converges.
        \item If $L > 1$, then the series $\sum a_n$ diverges.
        \item If $L = 1$, then the test is inconclusive.
    Proof of part 1: Let $L < 1$. Then, choose $\mu$ such that $L < \mu < 1$. Then for some $N$, $ \frac{a_{n+1}}{a_n}  < \mu$ for all $n > N$, then $a_{n} < \mu a_{n+1} < \mu^2 a_{n+2} < \ldots < \mu^j a_{n+j}$ for all $n > N$. So we have:
        \sum a_n &< \sum \frac{a_k}{\mu^k} \mu^n \\
        &= \frac{a_k}{\mu_k} \sum \mu^n
    Since $\sum \mu^n$ converges, then $\sum a_n$ converges.
    Let $a_n = \frac{k^2}{e^k}$ for some $k$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \left | \frac{a_{n+1}}{a_n} \right | &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{e^{k+1}} \cdot \frac{e^k}{k^2} \right | \\
        &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{k^2} \cdot \frac{e^k}{e^{k+1}} \right | \\
        &= \lim_{n \to \infty} \left | \frac{(k+1)^2}{k^2} \cdot \frac{1}{e} \right | \\
        &= \frac{1}{e} < 1
\subsection{Power Series}
\begin{definition}[Power Series]
    A power series is a series of the form:
        \sum_{n=0}^{\infty} c_n x^n = c_0 + c_1x + c_2x^2 + \ldots
    where $c_n$ are constants and $x$ is a variable. 
\begin{example}[Geometric Series]
    Let $c_n = 1$ for all $n$, then the series $\sum x^n$ is a geometric series:
        \sum x^n = 1 + x + x^2 + \ldots = \frac{1}{1-x}
\begin{definition}[Power Series at a Point]
    A power series at a point $a$ is a series of the form:
        \sum_{n=0}^{\infty} c_n (x-a)^n = c_0 + c_1(x-a) + c_2(x-a)^2 + \ldots
    where $c_n$ are constants and $x$ is a variable.
\paragraph{} A special convergent limit for the power series when $x = a$ is called the radius of convergence. This always converges to $c_0$.
    Let $a_n = \frac{x^n}{n^2}$ for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1} \cdot n^2}{x^n \cdot (n+1)^2} \right| \\
        &= \lim_{n \to \infty} \left| x \cdot \frac{n^2}{(n+1)^2} \right| \\
        &= \left| x \right| \lim_{n \to \infty} \left( \frac{n}{n+1} \right)^2 \\
        &= \left| x \right|
According to the ratio test, the series $\sum a_n$ converges if $\left| x \right| < 1$. And if $x=1$ it is a $p=2$ series which converges. And for $x=-1$ we can use the alternating series test to show that the series converges since $|a_{n+1}| \le |a_n|$ and $\lim_{n \to \infty} a_n = 0$.
    Let $a_n = \frac{x^n(1+5^n)}{n!}$  for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges. We have:
        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1}(1+5^{n+1})}{(n+1)!} \cdot \frac{n!}{x^n(1+5^n)} \right| \\
        &= \lim_{n \to \infty} \left| \frac{x}{1+n} \cdot \frac{1+5^{n+1}}{1+5^n} \right| \\
        \intertext{The fist part goes to 5, and the second part goes to 0. Thus, the limit is 0.}
        &= 0
According to the ratio test, the series $\sum a_n$ converges.
    Let $a_n = x^n n!$ for some $x$. We can use the ratio test to show that the series $\sum a_n$ converges only when $x=0$. We have:
        \lim_{n \to \infty} \left| \frac{a_{n+1}}{a_n} \right| &= \lim_{n \to \infty} \left| \frac{x^{n+1}(n+1)!}{x^n n!} \right| \\
        &= \lim_{n \to \infty} \left| x(n+1) \right| 
    Thus, the series converges only when $x=0$.

### 2025-02-08 MAT185 Lecture Notes/main.tex
feb 8

\begin{definition}[Column Space]
    Let $a_1, a_2, \ldots, {a_n} \in {^m\mathbb{R}}$, then:   
    \begin{equation}
        \text{Col}(A) = \text{Span}({a_1}, {a_2}, \ldots, {a_n})
    \end{equation}
    is the column space of $A$.
\end{definition}
Note that the span of set is a subspace of the vector space. Thus, the column space of a matrix is a subspace of $^m\mathbb{R}$. This is also call the set of images of the transformation. We can also note that matrix multiplication is is commutative with scalar multiplication. We can conduct proofs using such properties.
\begin{definition}[Rank]
    Let $A \in {^m\mathbb{R}^n}$, then:
    \begin{equation}
        \text{rank}(A) = \text{dim}(\text{Col}(A))
    \end{equation}
    is the rank of $A$.
\end{definition}
\end{document}

>>>>

\begin{definition}[Column Space]
    Let $a_1, a_2, \ldots, {a_n} \in {^m\mathbb{R}}$, then:   
    \begin{equation}
        \text{Col}(A) = \text{Span}({a_1}, {a_2}, \ldots, {a_n})
    \end{equation}
    is the column space of $A$.
\end{definition}
\begin{definition}[Row Space]
    Let $a_1, a_2, \ldots, {a_n} \in {^m\mathbb{R}}$, then:   
    \begin{equation}
        \text{Row}(A) = \text{Span}({a_1}, {a_2}, \ldots, {a_n})
    \end{equation}
    is the row space of $A$.
\end{definition}
Note that the span of set is a subspace of the vector space. Thus, the column space of a matrix is a subspace of $^m\mathbb{R}$. This is also call the set of images of the transformation. We can also note that matrix multiplication is is commutative with scalar multiplication. We can conduct proofs using such properties.
\begin{definition}[Rank]
    Let $A \in {^m\mathbb{R}^n}$, then:
    \begin{equation}
        \text{rank}(A) = \text{dim}(\text{Col}(A))
    \end{equation}
    is the rank of $A$.
\end{definition}
\begin{theorem}[Rank of RREF]
    Let $A \in {^m\mathbb{R}^n}$, then:
    \begin{equation}
        \text{rank}(A) = \text{rank}(\text{RREF}(A))
    \end{equation}
\end{theorem}
\begin{theorem}[Dimension of Row and Column Space]
    Let $A \in {^m\mathbb{R}^n}$, then:
    \begin{align}
        \text{dim}(\text{Row}(A)) &= \text{rank}(A) \\
        \text{dim}(\text{Col}(A)) &= \text{rank}(A)
    \end{align}
\end{theorem}
\begin{definition}[Matrix Multiplication]
    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
    \begin{equation}
        (AB)_{ij} = \sum_{k=1}^{n} a_{ik}b_{kj}
    \end{equation}
\end{definition}
\begin{definition}[Column View of Matrix Multiplication]
    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
    \begin{equation}
        AB = \begin{bmatrix} A\textbf{b}_1 & A\textbf{b}_2 & \ldots & A\textbf{b}_p \end{bmatrix}
    \end{equation}
\end{definition}
\begin{theorem}[Column Space of Matrix Multiplication]
    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
    \begin{equation}
        \text{Col}(AB) \subseteq \text{Col}(A)
    \end{equation}
    If $B$ is a invertible matrix, then also $\text{Col}(A) \subseteq \text{Col}(AB)$
\end{theorem}
\begin{definition}[Invertible]
    Let $A \in {^n\mathbb{R}^n}$, then $A$ is invertible if there exists a matrix $B \in {^n\mathbb{R}^n}$ such that:
    \begin{align}
        AB = BA = I
    \end{align}
    where $I$ is the identity matrix, and we denote $B = A^{-1}$, the inverse of $A$.
\end{definition}
\begin{proof}
    Let $A$, $B$ be matrices, so $\text{col}(AB) = \text{Span}(\text{columns of }AB)$. In the column view, we have $AB = A(\text{columns of }B) = \begin{bmatrix}
        A\textbf{b}_1 & A\textbf{b}_2 & \ldots & A\textbf{b}_p
    \end{bmatrix}$
    Each column of $AB$ is in $\text{Col}(A)$. So each column of $AB$ is a linear combination of the columns of $A$. Thus, $\text{Col}(AB) \subseteq \text{Col}(A)$.

    Consider the case where $B$ is invertible. Let $C = AB$, Then $CB^{-1} = A$. Since it is proven that $\text{col}(CB^{-1}) \subseteq \text{col}(C)$, hence,  $\text{col}(A) \subseteq \text{col}(AB)$.
\end{proof}
\begin{definition}[Null Space of Matrix Multiplication]
    Let $A \in {^m\mathbb{R}^n}$ and $B \in {^n\mathbb{R}^p}$, then:
    \begin{equation}
        \text{null}(AB) \supseteq \text{null}(B)
    \end{equation}
\end{definition}
\end{document}

### 2025-02-08 MSE160 Lecture Notes/main.tex
feb 8

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MSE 160 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Molecules and Materials, Winter 2024 \hfill MSE 160\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
\begin{center}
\textit{"In this class we are mostly understanding solids''} \\ - Prof. \textsc{Scott Ramsay}
\end{center}
\vspace{10pt}
\section{Mechanical Behavior}
\paragraph{Classes of Materials} In this class, we look at three classes of materials (non-exhaustive):
\begin{itemize}
    \item \textbf{Metal} held together with metallic bonds, typically \textbf{ductile} and \textbf{conductive}.
    \item \textbf{Ceramics}  (often metal oxides [excp: diamond]) held together via covalent \& ionic bonds, typically \textbf{brittle} and \textbf{insulating}.
    \item \textbf{Polymers} Molecules (often hydrocarbons) typically \textbf{ductile} and \textbf{insulating}
\end{itemize}
\paragraph{Engineering Stress} For normal stress, we know that:
\begin{equation}
    \sigma = \frac{F}{A_0}
\end{equation}
\paragraph{Engineering Strein} Also:
\begin{equation}
    \epsilon = \frac{\Delta l}{l_0}
\end{equation}
\paragraph{Young's Moduclus} For elastic deformation, $E$, is given, by Hooke's Law, as follows:
\begin{equation}
    \sigma = E \epsilon
\end{equation}
\paragraph{Tensile Test} We apply force as to the ends of a dogbone-sample, with $l_0$ being the gauge length and $A_0$ being the area of the cross-section at the middle. 
\paragraph{Tensile Strein} Maximum tensile strain on the engineeging stress-strain curve.
\subsection{Understanding Elastic Properties in terms of Atomic Configuration}
\paragraph{Atomic Configuration} We can understand the elastic properties of a material by looking at the atomic configuration. Skemetically, we can represent the atomic configuration as a spring system:
\begin{enumerate}
    \item \textbf{Intial - Before Loading} Atoms are in equilibrium, with the interatomic forces being balanced.
    \item \textbf{Loading} We apply a force to the material, causing the atoms to move from their equilibrium positions. The bond stretches and the atoms move further apart.
    \item \textbf{Unloading} We remove the force, causing the atoms to return to their equilibrium positions. 
\end{enumerate}
\paragraph{Atom Positions} Elastic modulus is dependent on the atomic interatomic bonding force. Thus, The elastic modulus is prootionaly to the slope of the interatomic force-seperation curve.
\paragraph{Force-Seperation Curve} The force-seperation curve is a plot of the force between two atoms as a function of the distance between them. The slope of the curve is proportional to the elastic modulus near the equilibrium position.

@@

\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}

>>>>

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MSE 160 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Molecules and Materials, Winter 2024 \hfill MSE 160\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\
\begin{center}
\textit{"In this class we are mostly understanding solids''} \\ - Prof. \textsc{Scott Ramsay}
\end{center}
% TODO: add a to r ratio table for different crystal structures/cubic structures
\vspace{10pt}
\section{Mechanical Behavior}
\paragraph{Classes of Materials} In this class, we look at three classes of materials (non-exhaustive):
\begin{itemize}
    \item \textbf{Metal} held together with metallic bonds, typically \textbf{ductile} and \textbf{conductive}.
    \item \textbf{Ceramics}  (often metal oxides [excp: diamond]) held together via covalent \& ionic bonds, typically \textbf{brittle} and \textbf{insulating}.
    \item \textbf{Polymers} Molecules (often hydrocarbons) typically \textbf{ductile} and \textbf{insulating}
\end{itemize}
\paragraph{Engineering Stress} For normal stress, we know that:
\begin{equation}
    \sigma = \frac{F}{A_0}
\end{equation}
\paragraph{Engineering Strein} Also:
\begin{equation}
    \epsilon = \frac{\Delta l}{l_0}
\end{equation}
\paragraph{Young's Moduclus} For elastic deformation, $E$, is given, by Hooke's Law, as follows:
\begin{equation}
    \sigma = E \epsilon
\end{equation}
\paragraph{Tensile Test} We apply force as to the ends of a dogbone-sample, with $l_0$ being the gauge length and $A_0$ being the area of the cross-section at the middle. 
\paragraph{Tensile Strein} Maximum tensile strain on the engineeging stress-strain curve.
\subsection{Understanding Elastic Properties in terms of Atomic Configuration}
\paragraph{Atomic Configuration} We can understand the elastic properties of a material by looking at the atomic configuration. Skemetically, we can represent the atomic configuration as a spring system:
\begin{enumerate}
    \item \textbf{Intial - Before Loading} Atoms are in equilibrium, with the interatomic forces being balanced.
    \item \textbf{Loading} We apply a force to the material, causing the atoms to move from their equilibrium positions. The bond stretches and the atoms move further apart.
    \item \textbf{Unloading} We remove the force, causing the atoms to return to their equilibrium positions. 
\end{enumerate}
\paragraph{Atom Positions} Elastic modulus is dependent on the atomic interatomic bonding force. Thus, The elastic modulus is prootionaly to the slope of the interatomic force-seperation curve.
\paragraph{Force-Seperation Curve} The force-seperation curve is a plot of the force between two atoms as a function of the distance between them. The slope of the curve is proportional to the elastic modulus near the equilibrium position.

@@

\section{Stress-Strain Behavior}
\paragraph{Uniformly Distributed Elongation} The elongation of a material is uniformly distributed within the gauge length. The initial values, referred to as the ``elastic region," are characterized as uniformly elastic. 
\paragraph{Non-Linear Elastic} There is also a non-linear region where the material remains elastic but is not uniformly distributed.
\paragraph{Localized Plastic Deformation} The material then enters the plastic region where the deformation is localized. The material is permanently deformed. This is characterized by the ultimate tensile strength (this is bad when it happens).
\paragraph{Necking} The material then necks, where the cross-sectional area decreases. The material is then strained until it breaks.
\paragraph{True Stress} The true stress is given by:
    \sigma = \frac{F}{A}
, where $A$ is the instantaneous cross-sectional area.
\paragraph{Ultimate Tensile Strength} The ultimate tensile strength is the maximum stress on the stress-strain curve.
\paragraph{Strengthening} Plastic deformation is made more difficult by the following mechanisms:
    \item \textbf{Strain Hardening} Preloading beyond the elastic region - upon plastic deformation, the material will have a higher yield strength.
    \item \textbf{For Metals and Alloys} The following mechanisms are used:

\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}
